Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-12-09 submit/fix.tex
fixed normalization bug in stan

\begin{abstract}
Through empirical modeling of its observed signal, the peak absolute magnitude of a Type~Ia supernova (SN~Ia) can be accurately determined,
making SNe~Ia  excellent distance indicators.  Improved modeling of SN~Ia colors and magnitudes can account for
further physical
diversity that is expected but not included
in current models, and so consequently can lead to more precise per-object distances with smaller systematic uncertainties.  In this article, we present   
an empirical model for SN~Ia peak colors with
\added{three color parameters and dependence on the equivalent widths of Ca~II and Si~II, the Si~II velocity,
and the light-curve shape.
This model is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the observable features and on the three color parameters.
The color changes allowed by two of the color parameters are similar to those allowed by the extinction model of  \citet{1999PASP..111...63F}. 
Correspondingly the sample has an effective value of $\langle R^F_{\mathit{eff}}\rangle \sim 2.40$, though with a broad distribution of 
total-to-selective extinctions. 
We detect the influence on color by a third parameter at
%--- table.py
$> 99\%$
%-----
confidence.
The third parameter has  unique color characteristics that distinguish it from 
those already  tracked by the spectral features, light-curve shape, and two-parameter dust models.}
\end{abstract}

@@

\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Parameters& $X={\hat{U}}$ &${\hat{B}}$&${\hat{V}}$&${\hat{R}}$&${\hat{I}}$\\ \hline
$\alpha_X$
& $0.0059^{+0.0007}_{-0.0008}$
& $0.0033^{+0.0006}_{-0.0007}$
& $0.0029^{+0.0006}_{-0.0006}$
& $0.0026^{+0.0005}_{-0.0005}$
& $0.0041^{+0.0004}_{-0.0005}$
\\
${\alpha_X/\alpha_{\hat{V}}-1}$
& $   1.0^{+   0.2}_{  -0.2}$
& $   0.1^{+   0.1}_{  -0.1}$
&\dots
& $  -0.1^{+   0.0}_{  -0.0}$
& $   0.4^{+   0.2}_{  -0.1}$
\\
$\beta_X$
& $ 0.032^{+ 0.006}_{-0.006}$
& $ 0.021^{+ 0.005}_{-0.005}$
& $ 0.018^{+ 0.005}_{-0.005}$
& $ 0.016^{+ 0.004}_{-0.004}$
& $ 0.008^{+ 0.004}_{-0.004}$
\\
${\beta_X/\beta_{\hat{V}}-1}$
& $  0.79^{+  0.29}_{ -0.19}$
& $  0.16^{+  0.11}_{ -0.08}$
&\dots
& $ -0.12^{+  0.04}_{ -0.04}$
& $ -0.58^{+  0.11}_{ -0.17}$
\\
$\eta_X$
& $0.0022^{+0.0011}_{-0.0011}$
& $0.0024^{+0.0009}_{-0.0010}$
& $0.0027^{+0.0008}_{-0.0008}$
& $0.0025^{+0.0007}_{-0.0007}$
& $0.0018^{+0.0006}_{-0.0006}$
\\
${\eta_X/\eta_{\hat{V}}-1}$
& $ -0.19^{+  0.17}_{ -0.28}$
& $ -0.12^{+  0.09}_{ -0.15}$
&\dots
& $ -0.06^{+  0.06}_{ -0.04}$
& $ -0.33^{+  0.10}_{ -0.10}$
\\
$\zeta_X$
& $ -0.06^{+  0.05}_{ -0.04}$
& $ -0.09^{+  0.04}_{ -0.04}$
& $ -0.12^{+  0.04}_{ -0.04}$
& $ -0.09^{+  0.03}_{ -0.03}$
& $ -0.14^{+  0.03}_{ -0.03}$
\\
${\zeta_X/\zeta_{\hat{V}}-1}$
& $ -0.45^{+  0.20}_{ -0.33}$
& $ -0.26^{+  0.11}_{ -0.18}$
&\dots
& $ -0.20^{+  0.05}_{ -0.07}$
& $  0.22^{+  0.23}_{ -0.13}$
\\
$\gamma^0_X$
& $ 68.29^{+  2.97}_{ -2.92}$
& $ 54.90^{+  2.49}_{ -2.56}$
& $ 40.89^{+  2.38}_{ -2.47}$
& $ 30.95^{+  2.13}_{ -2.25}$
& $ 22.34^{+  2.05}_{ -2.16}$
\\
${\gamma^0_X/\gamma^0_{\hat{V}}-1}$
& $  0.67^{+  0.05}_{ -0.05}$
& $  0.34^{+  0.03}_{ -0.03}$
&\dots
& $ -0.24^{+  0.01}_{ -0.01}$
& $ -0.45^{+  0.03}_{ -0.03}$
\\
$\gamma^1_X$
& $-10.15^{+  4.45}_{ -4.49}$
& $ -6.81^{+  3.88}_{ -3.82}$
& $-15.06^{+  3.42}_{ -3.39}$
& $-14.88^{+  3.01}_{ -3.00}$
& $-15.91^{+  2.90}_{ -2.86}$
\\
${\gamma^1_X/\gamma^1_{\hat{V}}-1}$
& $ -0.32^{+  0.16}_{ -0.21}$
& $ -0.55^{+  0.13}_{ -0.20}$
&\dots
& $ -0.01^{+  0.05}_{ -0.04}$
& $  0.06^{+  0.14}_{ -0.10}$
\\
$\sigma_p \phi_X$
& $ 0.024^{+ 0.009}_{-0.009}$
& $-0.018^{+ 0.007}_{-0.007}$
& $-0.027^{+ 0.006}_{-0.006}$
& $-0.010^{+ 0.006}_{-0.006}$
& $ 0.022^{+ 0.006}_{-0.006}$
\\
${\phi_X/\phi_{\hat{V}}-1}$
& $-1.898^{+ 0.420}_{-0.611}$
& $-0.333^{+ 0.156}_{-0.181}$
&\dots
& $-0.609^{+ 0.109}_{-0.163}$
& $-1.834^{+ 0.329}_{-0.486}$
\\
\hline
\end{tabular}
\caption{68\% credible intervals for the global fit parameters of Model~III in \S\ref{modelIII:sec}.\label{global3:tab}}
\end{table}

@@

The transformation between the $\vec{\gamma}$  vectors and the   \citetalias{1999PASP..111...63F} vector given in
Eq.~\ref{trans_I:eqn}, is satisfied by the transformation matrix
\begin{equation}
M=
\begin{pmatrix}
\begin{array}{rr}
39.7^{+2.4}_{-2.5} & 16.6^{+0.6}_{-0.6}  \\
-14.7^{+3.4}_{-3.4} & 3.7^{+1.0}_{-0.9}  
\end{array}
\end{pmatrix} 
\end{equation}
and the two residual vectors
\begin{align}
\begin{split}
\vec{\epsilon}_{\gamma^0} &=\left(1.0^{+0.4}_{-0.4} , -1.2^{+0.4}_{-0.4} , -0.8^{+0.4}_{-0.4} , 0.7^{+0.2}_{-0.2} , 0.4^{+0.6}_{-0.6}\right), \\
\vec{\epsilon}_{\gamma^1} & =\left(-2.7^{+0.6}_{-0.6} , 4.3^{+0.6}_{-0.6} , -0.8^{+0.6}_{-0.6} , 1.2^{+0.3}_{-0.3} , -2.6^{+0.9}_{-0.9}\right).
\end{split}
\label{res_3:eqn}
\end{align}
This corresponds to a quadratic fraction of $0.9996^{+0.0002}_{-0.0003}$ of $\vec{\gamma}^0$ that projects onto the plane of
color changes allowed by the \citetalias{1999PASP..111...63F} dust-extinction model
as described by
$\vec{a}$ and $\vec{b}$. The quadratic fraction of $\vec{\gamma}^1$ projected
onto this plane is
$0.959^{+0.015}_{-0.029}$.
The   $\vec{\gamma}^0$  vector points in the $\langle R^F_{\mathit{eff}}\rangle=2.40^{+0.15}_{-0.15}$ dust-extinction direction.
This value is consistent with the $R_V=2.5$ found in the supernova samples of \citet{2011ApJ...729...55F}, and the
center  of the range of per-supernova $R^F$ values found in the samples of \citet{2014ApJ...789...32B, 2015MNRAS.453.3300A}.
Based on these facts, the supernova ``extrinsic'' colors $\vec{\gamma}^0 g_0 + \vec{\gamma}^1 g_1$  of our model are attributed to dust extinction.
This association is further discussed when $g_0$ and $g_1$ are considered in \S\ref{g0g1:sec}.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{../output_fix3_x1/plane0.pdf}
   \includegraphics[width=2.95in]{../output_fix3_x1/plane1.pdf}
   \caption{
   Visualization of how supernova magnitudes can vary in Model~III and the dust extinction of \citetalias{1999PASP..111...63F}.
   See Figure~\ref{plane:fig}.
   The combination $\vec{a}+\vec{b}/2.40$ is shown in the dotted red
   line: it is almost perfectly superimposed on $\vec{\gamma}^0$.
   \label{plane3:fig}}
\end{figure}

@@

The  new intrinsic $p$ parameter has a significant influence on color.  The magnitude of its color effect is non-zero,
with $\sigma_p = 0.049^{+ 0.004}_{-0.003}$~mag.  The smallest value of $\sigma_p$
over all Monte Carlo chains is 0.039~mag.
These numbers are significantly larger than the median data color uncertainty of 0.010~mag (the maximum is 0.042~mag).
The evaluation of 100 simulated data sets with no third color parameter produces no median $\sigma_p$'s  as
high as that of the data, so we assign a $>99$\% confidence in the detection of this third color term.

The direction of $\vec{\phi}$ is not aligned with the plane defined by
the $\vec{\gamma}$ vectors (nor of the  \citetalias{1999PASP..111...63F} dust-extinction model); a quadratic fraction of
$0.053^{+0.102}_{-0.040}$ of 
 $\vec{\phi}$ lies in the $\vec{\gamma}^0$--$\vec{\gamma}^1$ plane.

@@

The Pearson correlation coefficients of the per-supernova parameters are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), A_{p,V}) =\\
\begin{pmatrix}
\begin{array}{rrrrrrrr}
\Delta & -0.07^{+0.06}_{-0.06} & -0.11^{+0.08}_{-0.07} & -0.17^{+0.06}_{-0.06} & 0.19^{+0.08}_{-0.08} & 0.08^{+0.07}_{-0.06} & 0.23^{+0.08}_{-0.08} & -0.13^{+0.06}_{-0.06} \\
\ldots  & EW_{Ca} & 0.09^{+0.03}_{-0.03} & -0.26^{+0.03}_{-0.03} & 0.07^{+0.03}_{-0.03} & -0.11^{+0.05}_{-0.05} & -0.18^{+0.06}_{-0.06} & -0.00^{+0.08}_{-0.07} \\
\ldots & \ldots &EW_{Si} & -0.12^{+0.03}_{-0.03} & -0.84^{+0.01}_{-0.01} & -0.13^{+0.04}_{-0.04} & 0.27^{+0.08}_{-0.09} & -0.00^{+0.08}_{-0.08} \\
\ldots  & \ldots  & \ldots  &  \lambda_{Si} & 0.14^{+0.03}_{-0.03} & 0.01^{+0.05}_{-0.05} & 0.00^{+0.06}_{-0.06} & 0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  &\ldots & \ldots & x_1 & 0.13^{+0.04}_{-0.04} & -0.20^{+0.09}_{-0.09} & 0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  & \ldots  &\ldots  & \ldots  & E_{\gamma^0}({\hat{B}}-{\hat{V}})& 0.04^{+0.06}_{-0.06} & -0.00^{+0.07}_{-0.08} \\
\ldots  & \ldots &\ldots  & \ldots  & \ldots & \ldots  &E_{\gamma^1}({\hat{B}}-{\hat{V}})  & 0.00^{+0.07}_{-0.07} \\
\ldots  & \ldots & \ldots & \ldots  & \ldots  & \ldots  &\ldots  & A_{p,V} \\
\end{array}
\end{pmatrix}.
\label{corr:eqn}
\end{multline}
To aid readability, only the upper triangular part of the otherwise symmetric matrix is shown and the corresponding feature label is given on
the diagonal.

@@

\subsubsection{Results for the Latent Extrinsic Parameters $g_0$, $g_1$}
\label{g0g1:sec}
The vector
$\vec{\gamma}^0$ describes color changes for an effective $\langle R^F_{\mathit{eff}}\rangle=2.40$ dust, as shown 
in  \S\ref{results3gamma:sec}.
Hence, the per-supernova parameter $\gamma^0_{\hat{V}} g_0$ is associated with the $\hat{V}$-band
extinction for a constant $R^F=2.40$.}
Figure~\ref{k0_med:fig} shows the histogram of
per-supernova
median values
from the Monte Carlo chains
and the  stack of the posteriors of all supernovae  for $\gamma^0_{\hat{V}} g_0$ 
relative to that of an arbitrary supernova  $\gamma^0_{\hat{V}} g_0|_0$.
The distributions are non-Gaussian, having a sharp rise in the blue and an extended tail in the red.  This is consistent
with
simulations based on expected dust distributions
within galaxies and the distribution of galaxy orientations with respect to the observer,
which determine column densities along the lines of sight toward supernovae that are consistent with observations
\citep{1998ApJ...502..177H, 2007ApJ...659..122J}.  
We emphasize that unlike other analyses,
this distribution is given by the data, and so is not dependent on any knowledge or prior
for the distribution of dust extinction the SN population is expected to suffer.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{../output_fix3_x1/deltagamma0_med.pdf}
   \includegraphics[width=2.8in]{../output_fix3_x1/deltagamma1_med.pdf}
      \caption{
      Normalized stack of the posteriors of all supernovae  and histogram 
      of  per-supernova 
median values of: (left) $\gamma^0_{\hat{V}} g_0-\gamma^0_{\hat{V}} g_0|_0$, which is associated with the $A^F_V$ of an $\langle R^F_{\mathit{eff}}\rangle \sim 2.4$ dust model;
(right) $\gamma^1_{\hat{V}} g_1-\gamma^1_{\hat{V}} g_1|_0$,
which  is associated with extinction corrections due to deviations away from the canonical $\langle R^F_{\mathit{eff}}\rangle$ value.
   \label{k0_med:fig}}
\end{figure}

\added{
An instructive way to consider the  $\{g_0, g_1\}$ parameters is
to transform them to $\{A_V^F, E^F(B-V)\}$ (modulo additive constants)   using the
matrix $M$ in  Eq.~\ref{trans_I:eqn}.
A plot showing the expected values and 68\% credible intervals of these parameters
for our supernova
sample is shown in Figure~\ref{kk:fig}. 
For reference, a line that represents $\langle R^F_{\mathit{eff}}\rangle=2.4$ is overplotted; points above
the line have larger $R^F$, points below have smaller $R^F$.  The majority of supernovae lie within a narrow range above the line,
while the remaining fraction fall in a broader range below the line.
Comparison with  $\gamma^1_{\hat{V}} g_1-\gamma^1_{\hat{V}} g_1|_0$ of the right plot of Figure~\ref{k0_med:fig}
shows that the 
negative tail corresponds to smaller $R^F$, and the sharper positive edge to larger $R^F$.
These findings are qualitatively consistent with previous results:
using fixed dust-extinction models,
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
Our model cannot provide the range of $R^F$ in our sample nor per-supernova determinations of $R^F$ 
without further assumptions.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{../output_fix3_x1/avebv_synth.pdf}
      \caption{
      Expected values and 68\% credible intervals of effective $E^F(B-V)$ and $A_V^F$ after transformation from our model $g_0$ and $g_1$ parameters, for the supernova in our sample.
      Overplotted is a line with the slope expected for $R^F=2.4$.
   \label{kk:fig}}
\end{figure}

@@

\subsubsection{Results for the Intrinsic Parameter $p$}
\label{p:sec}
The model assumes that $p$ has a Normal distribution, which is shown in Figure~\ref{ebv3:fig} as the 
stack of the posteriors of all supernovae of the external color-excess $E_{\gamma}({\hat{B}}-{\hat{V}})$.
Also shown for comparison is the distribution of
the extrinsic  color-excess $E_\gamma({\hat{B}}-{\hat{V}}) \equiv (\gamma^0_{\hat{B}} -\gamma^0_{\hat{V}}) g_0
+ (\gamma^1_{\hat{B}} -\gamma^1_{\hat{V}}) g_1$.
The standard deviations of  $E_\gamma({\hat{B}}-{\hat{V}})$ and $E_p({\hat{B}}-{\hat{V}})$ are
%-----
0.085
and 0.011
%-----
mag respectively.
The latter is comparable to the typical measurement uncertainty.
In $\hat{B}-\hat{V}$, 
the extrinsic color variation is much larger than that of the intrinsic and its effect per-supernova would be difficult to discern.

>>>>

\begin{abstract}
Through empirical modeling of its observed signal, the peak absolute magnitude of a Type~Ia supernova (SN~Ia) can be accurately determined,
making SNe~Ia  excellent distance indicators.  Improved modeling of SN~Ia colors and magnitudes can account for
further physical
diversity that is expected but not included
in current models, and so consequently can lead to more precise per-object distances with smaller systematic uncertainties.  In this article, we present   
an empirical model for SN~Ia peak colors with
\added{three color parameters and dependence on the equivalent widths of Ca~II and Si~II, the Si~II velocity,
and the light-curve shape.
This model is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the observable features and on the three color parameters.
The color changes allowed by two of the color parameters are similar to those allowed by the extinction model of  \citet{1999PASP..111...63F}. 
Correspondingly the sample has an effective value of $\langle R^F_{\mathit{eff}}\rangle \sim 2.39$, though with a broad distribution of 
total-to-selective extinctions. 
We detect the influence on color by a third parameter at
%--- table.py
$> 99\%$
%-----
confidence.
The third parameter has  unique color characteristics that distinguish it from 
those already  tracked by the spectral features, light-curve shape, and two-parameter dust models.}
\end{abstract}

@@

\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Parameters& $X={\hat{U}}$ &${\hat{B}}$&${\hat{V}}$&${\hat{R}}$&${\hat{I}}$\\ \hline
$\alpha_X$
& $0.0059^{+0.0007}_{-0.0008}$
& $0.0032^{+0.0006}_{-0.0007}$
& $0.0029^{+0.0006}_{-0.0006}$
& $0.0026^{+0.0005}_{-0.0005}$
& $0.0041^{+0.0004}_{-0.0004}$
\\
${\alpha_X/\alpha_{\hat{V}}-1}$
& $   1.0^{+   0.2}_{  -0.2}$
& $   0.1^{+   0.1}_{  -0.1}$
& \dots
& $  -0.1^{+   0.0}_{  -0.0}$
& $   0.4^{+   0.2}_{  -0.1}$
\\
$\beta_X$
& $ 0.032^{+ 0.006}_{-0.006}$
& $ 0.021^{+ 0.005}_{-0.005}$
& $ 0.018^{+ 0.005}_{-0.005}$
& $ 0.016^{+ 0.004}_{-0.004}$
& $ 0.007^{+ 0.004}_{-0.004}$
\\
${\beta_X/\beta_{\hat{V}}-1}$
& $  0.80^{+  0.29}_{ -0.19}$
& $  0.17^{+  0.11}_{ -0.08}$
& \dots
& $ -0.13^{+  0.04}_{ -0.04}$
& $ -0.59^{+  0.12}_{ -0.18}$
\\
$\eta_X$
& $0.0021^{+0.0011}_{-0.0012}$
& $0.0023^{+0.0009}_{-0.0010}$
& $0.0026^{+0.0008}_{-0.0008}$
& $0.0025^{+0.0007}_{-0.0007}$
& $0.0018^{+0.0006}_{-0.0006}$
\\
${\eta_X/\eta_{\hat{V}}-1}$
& $ -0.22^{+  0.19}_{ -0.33}$
& $ -0.13^{+  0.10}_{ -0.17}$
& \dots
& $ -0.06^{+  0.07}_{ -0.05}$
& $ -0.32^{+  0.10}_{ -0.11}$
\\
$\zeta_X$
& $ -0.06^{+  0.05}_{ -0.04}$
& $ -0.08^{+  0.04}_{ -0.04}$
& $ -0.11^{+  0.04}_{ -0.04}$
& $ -0.09^{+  0.04}_{ -0.03}$
& $ -0.14^{+  0.03}_{ -0.03}$
\\
${\zeta_X/\zeta_{\hat{V}}-1}$
& $ -0.47^{+  0.20}_{ -0.35}$
& $ -0.27^{+  0.11}_{ -0.19}$
& \dots
& $ -0.20^{+  0.05}_{ -0.07}$
& $  0.24^{+  0.25}_{ -0.14}$
\\
$\gamma^0_X$
& $ 68.30^{+  2.98}_{ -2.99}$
& $ 54.82^{+  2.57}_{ -2.64}$
& $ 40.81^{+  2.45}_{ -2.56}$
& $ 30.89^{+  2.19}_{ -2.28}$
& $ 22.34^{+  2.11}_{ -2.17}$
\\
${\gamma^0_X/\gamma^0_{\hat{V}}-1}$
& $  0.68^{+  0.05}_{ -0.05}$
& $  0.34^{+  0.03}_{ -0.03}$
& \dots
& $ -0.24^{+  0.01}_{ -0.01}$
& $ -0.45^{+  0.03}_{ -0.03}$
\\
$\gamma^1_X$
& $ -9.92^{+  4.29}_{ -4.39}$
& $ -6.59^{+  3.87}_{ -3.90}$
& $-14.86^{+  3.51}_{ -3.43}$
& $-14.70^{+  3.03}_{ -3.00}$
& $-15.74^{+  2.82}_{ -2.80}$
\\
${\gamma^1_X/\gamma^1_{\hat{V}}-1}$
& $ -0.33^{+  0.16}_{ -0.21}$
& $ -0.56^{+  0.14}_{ -0.21}$
& \dots
& $ -0.01^{+  0.06}_{ -0.04}$
& $  0.06^{+  0.15}_{ -0.11}$
\\
$\sigma_p \phi_X$
& $ 0.020^{+ 0.014}_{-0.016}$
& $-0.022^{+ 0.012}_{-0.013}$
& $-0.031^{+ 0.010}_{-0.012}$
& $-0.014^{+ 0.009}_{-0.010}$
& $ 0.020^{+ 0.009}_{-0.010}$
\\
${\phi_X/\phi_{\hat{V}}-1}$
& $-1.658^{+ 0.559}_{-0.960}$
& $-0.278^{+ 0.168}_{-0.245}$
& \dots
& $-0.555^{+ 0.123}_{-0.219}$
& $-1.654^{+ 0.394}_{-0.701}$
\\
\hline
\end{tabular}
\caption{68\% credible intervals for the global fit parameters of Model~III in \S\ref{modelIII:sec}.\label{global3:tab}}
\end{table}

@@

The transformation between the $\vec{\gamma}$  vectors and the   \citetalias{1999PASP..111...63F} vector given in
Eq.~\ref{trans_I:eqn}, is satisfied by the transformation matrix
\begin{equation}
M=
\begin{pmatrix}
\begin{array}{rr}
39.7^{+2.4}_{-2.5} & 16.6^{+0.7}_{-0.6} \\
-14.5^{+3.4}_{-3.4} & 3.7^{+1.0}_{-1.0}
\end{array}
\end{pmatrix} 
\end{equation}
and the two residual vectors
\begin{align}
\begin{split}
\vec{\epsilon}_{\gamma^0} &=\left(1.1^{+0.4}_{-0.4} , -1.2^{+0.4}_{-0.4} , -0.8^{+0.4}_{-0.4} , 0.7^{+0.2}_{-0.2} , 0.4^{+0.6}_{-0.6}\right), \\
\vec{\epsilon}_{\gamma^1} & =\left(-2.8^{+0.6}_{-0.6} , 4.3^{+0.6}_{-0.6} , -0.8^{+0.6}_{-0.6} , 1.2^{+0.3}_{-0.3} , -2.6^{+0.9}_{-0.9}\right).
\end{split}
\label{res_3:eqn}
\end{align}
This corresponds to a quadratic fraction of $0.9996^{+0.0002}_{-0.0003}$ of $\vec{\gamma}^0$ that projects onto the plane of
color changes allowed by the \citetalias{1999PASP..111...63F} dust-extinction model
as described by
$\vec{a}$ and $\vec{b}$. The quadratic fraction of $\vec{\gamma}^1$ projected
onto this plane is
$0.959^{+0.015}_{-0.029}$.
The   $\vec{\gamma}^0$  vector points in the $\langle R^F_{\mathit{eff}}\rangle=2.39^{+0.15}_{-0.16}$ dust-extinction direction.
This value is consistent with the $R_V=2.5$ found in the supernova samples of \citet{2011ApJ...729...55F}, and the
center  of the range of per-supernova $R^F$ values found in the samples of \citet{2014ApJ...789...32B, 2015MNRAS.453.3300A}.
Based on these facts, the supernova ``extrinsic'' colors $\vec{\gamma}^0 g_0 + \vec{\gamma}^1 g_1$  of our model are attributed to dust extinction.
This association is further discussed when $g_0$ and $g_1$ are considered in \S\ref{g0g1:sec}.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{../output_fix3_x1/plane0.pdf}
   \includegraphics[width=2.95in]{../output_fix3_x1/plane1.pdf}
   \caption{
   Visualization of how supernova magnitudes can vary in Model~III and the dust extinction of \citetalias{1999PASP..111...63F}.
   See Figure~\ref{plane:fig}.
   The combination $\vec{a}+\vec{b}/2.39$ is shown in the dotted red
   line: it is almost perfectly superimposed on $\vec{\gamma}^0$.
   \label{plane3:fig}}
\end{figure}

@@

The  new intrinsic $p$ parameter has a significant influence on color.  The magnitude of its color effect is non-zero,
with $\sigma_p = 0.053^{+ 0.010}_{-0.005}$~mag.  The smallest value of $\sigma_p$ % new
over all Monte Carlo chains is 0.040~mag.
These numbers are significantly larger than the median data color uncertainty of 0.010~mag (the maximum is 0.042~mag).
The evaluation of 100 simulated data sets with no third color parameter produces no median $\sigma_p$'s  as
high as that of the data, so we assign a $>99$\% confidence in the detection of this third color term.

The direction of $\vec{\phi}$ is not aligned with the plane defined by
the $\vec{\gamma}$ vectors (nor of the  \citetalias{1999PASP..111...63F} dust-extinction model); a quadratic fraction of
$0.130^{+0.262}_{-0.104}$ of  %new
 $\vec{\phi}$ lies in the $\vec{\gamma}^0$--$\vec{\gamma}^1$ plane.

@@

The Pearson correlation coefficients of the per-supernova parameters are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), A_{p,V}) =\\
\begin{pmatrix}
\begin{array}{rrrrrrrr}
\Delta& -0.08^{+0.06}_{-0.06} & -0.11^{+0.08}_{-0.07} & -0.17^{+0.06}_{-0.06} & 0.19^{+0.08}_{-0.09} & 0.07^{+0.07}_{-0.06} & 0.23^{+0.08}_{-0.08} & 0.12^{+0.06}_{-0.06} \\
\ldots  & EW_{Ca} & 0.10^{+0.03}_{-0.03} & -0.25^{+0.03}_{-0.03} & 0.07^{+0.03}_{-0.03} & -0.10^{+0.05}_{-0.05} & -0.18^{+0.06}_{-0.06} & 0.00^{+0.08}_{-0.08} \\
\ldots & \ldots &EW_{Si}  & -0.12^{+0.03}_{-0.03} & -0.84^{+0.01}_{-0.01} & -0.13^{+0.04}_{-0.04} & 0.27^{+0.08}_{-0.09} & -0.00^{+0.08}_{-0.07} \\
\ldots  & \ldots  & \ldots  &  \lambda_{Si}& 0.14^{+0.03}_{-0.03} & 0.01^{+0.06}_{-0.05} & 0.00^{+0.06}_{-0.06} & -0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  &\ldots & \ldots & x_1 & 0.13^{+0.04}_{-0.04} & -0.20^{+0.09}_{-0.09} & -0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  & \ldots  &\ldots  & \ldots  & E_{\gamma^0}({\hat{B}}-{\hat{V}})& 0.04^{+0.06}_{-0.06} & 0.00^{+0.08}_{-0.08} \\
\ldots  & \ldots &\ldots  & \ldots  & \ldots & \ldots  &E_{\gamma^1}({\hat{B}}-{\hat{V}})  & -0.00^{+0.08}_{-0.08} \\
\ldots  & \ldots & \ldots & \ldots  & \ldots  & \ldots  &\ldots  & A_{p,V} \\
\end{array}
\end{pmatrix}.
\label{corr:eqn}
\end{multline}
To aid readability, only the upper triangular part of the otherwise symmetric matrix is shown and the corresponding feature label is given on
the diagonal.

@@

\subsubsection{Results for the Latent Extrinsic Parameters $g_0$, $g_1$}
\label{g0g1:sec}
The vector
$\vec{\gamma}^0$ describes color changes for an effective $\langle R^F_{\mathit{eff}}\rangle=2.39$ dust, as shown 
in  \S\ref{results3gamma:sec}.
Hence, the per-supernova parameter $\gamma^0_{\hat{V}} g_0$ is associated with the $\hat{V}$-band
extinction for a constant $R^F=2.39$.}
Figure~\ref{k0_med:fig} shows the histogram of
per-supernova
median values
from the Monte Carlo chains
and the  stack of the posteriors of all supernovae  for $\gamma^0_{\hat{V}} g_0$ 
relative to that of an arbitrary supernova  $\gamma^0_{\hat{V}} g_0|_0$.
The distributions are non-Gaussian, having a sharp rise in the blue and an extended tail in the red.  This is consistent
with
simulations based on expected dust distributions
within galaxies and the distribution of galaxy orientations with respect to the observer,
which determine column densities along the lines of sight toward supernovae that are consistent with observations
\citep{1998ApJ...502..177H, 2007ApJ...659..122J}.  
We emphasize that unlike other analyses,
this distribution is given by the data, and so is not dependent on any knowledge or prior
for the distribution of dust extinction the SN population is expected to suffer.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{../output_fix3_x1/deltagamma0_med.pdf}
   \includegraphics[width=2.8in]{../output_fix3_x1/deltagamma1_med.pdf}
      \caption{
      Normalized stack of the posteriors of all supernovae  and histogram 
      of  per-supernova 
median values of: (left) $\gamma^0_{\hat{V}} g_0-\gamma^0_{\hat{V}} g_0|_0$, which is associated with the $A^F_V$ of an $\langle R^F_{\mathit{eff}}\rangle \sim 2.39$ dust model;
(right) $\gamma^1_{\hat{V}} g_1-\gamma^1_{\hat{V}} g_1|_0$,
which  is associated with extinction corrections due to deviations away from the canonical $\langle R^F_{\mathit{eff}}\rangle$ value.
   \label{k0_med:fig}}
\end{figure}

\added{
An instructive way to consider the  $\{g_0, g_1\}$ parameters is
to transform them to $\{A_V^F, E^F(B-V)\}$ (modulo additive constants)   using the
matrix $M$ in  Eq.~\ref{trans_I:eqn}.
A plot showing the expected values and 68\% credible intervals of these parameters
for our supernova
sample is shown in Figure~\ref{kk:fig}. 
For reference, a line that represents $\langle R^F_{\mathit{eff}}\rangle=2.39$ is overplotted; points above
the line have larger $R^F$, points below have smaller $R^F$.  The majority of supernovae lie within a narrow range above the line,
while the remaining fraction fall in a broader range below the line.
Comparison with  $\gamma^1_{\hat{V}} g_1-\gamma^1_{\hat{V}} g_1|_0$ of the right plot of Figure~\ref{k0_med:fig}
shows that the 
negative tail corresponds to smaller $R^F$, and the sharper positive edge to larger $R^F$.
These findings are qualitatively consistent with previous results:
using fixed dust-extinction models,
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
Our model cannot provide the range of $R^F$ in our sample nor per-supernova determinations of $R^F$ 
without further assumptions.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{../output_fix3_x1/avebv_synth.pdf}
      \caption{
      Expected values and 68\% credible intervals of effective $E^F(B-V)$ and $A_V^F$ after transformation from our model $g_0$ and $g_1$ parameters, for the supernova in our sample.
      Overplotted is a line with the slope expected for $R^F=2.39$.
   \label{kk:fig}}
\end{figure}

@@

\subsubsection{Results for the Intrinsic Parameter $p$}
\label{p:sec}
The model assumes that $p$ has a Normal distribution, which is shown in Figure~\ref{ebv3:fig} as the 
stack of the posteriors of all supernovae of the external color-excess $E_{\gamma}({\hat{B}}-{\hat{V}})$.
Also shown for comparison is the distribution of
the extrinsic  color-excess $E_\gamma({\hat{B}}-{\hat{V}}) \equiv (\gamma^0_{\hat{B}} -\gamma^0_{\hat{V}}) g_0
+ (\gamma^1_{\hat{B}} -\gamma^1_{\hat{V}}) g_1$.
The standard deviations of  $E_\gamma({\hat{B}}-{\hat{V}})$ and $E_p({\hat{B}}-{\hat{V}})$ are
%-----
0.085 %new
and 0.010
%-----
mag respectively.
The latter is comparable to the typical measurement uncertainty.
In $\hat{B}-\hat{V}$, 
the extrinsic color variation is much larger than that of the intrinsic and its effect per-supernova would be difficult to discern.
